Self-tunable approximated explicit MPC: Heat exchanger implementation and analysys

### 2023-01-27 paper/elsarticle/main.tex
Review: Conclusions

@@

@@

\subsection{Self-tunable technique for systems with asymmetric behavior}
\label{sec:self_tunable_rho_scaling}
\noindent TODO: co ak nepriama umera? \\
TODO: remark na nelinearnu interpolaciu \\
TODO: zovseobecnit na viac usekov intervalu? \\

@@

The plant on which the control was implemented and analyzed is a laboratory counter-current liquid-to-liquid plate heat exchanger Armfield Process Plant Trainer PCT23~\cite{pct23}, see Figure~\ref{fig:HE}. The cold feed as well as heating medium are transported to the heat exchanger by two peristaltic pumps. The flow rate of the feed of the cold medium (water) is constant, while the aim of control is to track the reference value of the cold medium temperature. Therefore, the controlled variable is the feed temperature $T$ at the outlet of the heat exchanger. The associated manipulated variable is the voltage $U$ corresponding to the power of the pump feeding the heat exchanger by the hot medium (water). The voltage is within the range of [$0-5$]\,V normalized into the relative values in percentage. The manipulated variable is constrained from 20\,\% to 100\,\% to respect the physical limitations of the operating conditions. As the heat exchange is a nonlinear and asymmetric process~\cite{Liptak}, this heat exchanger represents a suitable candidate for the presented controller tuning strategy.

@@

\section{Conclusion}
\label{sec:conclusion}

This paper deals with the experimental implementation of tunable approximated explicit model predictive control and provides a technique for self-tuning controller design. Based on the current reference value, the tuning parameter is scaled using linear interpolation. The previously published paper related to the self-tunable explicit MPC suggested tuning based on the distance of the reference value from the steady state. This paper presents a novel idea of tuning based on the size of reference step change. The tuning procedures aim to compensate for the nonlinear behavior of the controlled system. The tuning parameter is updated when the reference changes. It is calculated as the ratio between the size of the reference change and the maximal possible size of the reference change, which is specified before operation. Another novel contribution comes with splitting the interval of tuning parameter into two ranges, while both ranges are assigned to different operating conditions. The proposed method is implemented on a laboratory heat exchanger with nonlinear and asymmetric behavior. The non-symmetry makes the plant a suitable candidate for splitting the interval of the tuning parameter. The decision criterium is negativity or positivity of reference change. When the reference changed upwards, the control input was tuned in the first part of the interval and approached the boundary controller associated with the lower bound on the penalty matrix. On the contrary, when the reference changed downwards, the control input was tuned to approach the control input from the boundary controller with the upper bound on the penalty matrix. To investigate the control results properly, the control performance was evaluated quantitatively. Compared to control with only one controller, the self-tunable technique helped to decrease the maximal overshoots/undershoots, integral square error, and settling time.

\section*{Acknowledgements}

>>>>

\textcolor{red}{[ TODO: Please, make consistent notation of equations: Eq-based vs. non-Eq notation. ]}

@@

\textcolor{red}{[ TODO: We need a well-elaborated introduction/review to strongly motivate our control approach. Please, take into account the industrial-oriented reviewers, mostly experts in the field of the chemical engineering (heat exchangers and energy balances). ]}

@@

\subsection{Self-tunable technique for systems with asymmetric behavior}
\label{sec:self_tunable_rho_scaling}

\noindent TODO: co ak nepriama umera? \\
TODO: remark na nelinearnu interpolaciu \\
TODO: zovseobecnit na viac usekov intervalu? \\

@@

The plant on which the control was implemented and analyzed is a laboratory-scaled counter-current liquid-to-liquid plate heat exchanger Armfield Process Plant Trainer PCT23~\cite{pct23}, see Figure~\ref{fig:HE}. The cold feed as well as heating medium are transported to the heat exchanger by two peristaltic pumps. The flow rate of the feed of the cold medium (water) is constant, while the aim of control is to track the reference value of the cold medium temperature. Therefore, the controlled variable is the feed temperature $T$ at the outlet of the heat exchanger. The associated manipulated variable is the voltage $U$ corresponding to the power of the pump feeding the heat exchanger by the hot medium (water). The voltage is within the range of [$0-5$]\,V normalized into the relative values in percentage. The manipulated variable is constrained from 20\,\% to 100\,\% to respect the physical limitations of the operating conditions. As the heat exchange is a nonlinear and asymmetric process~\cite{Liptak}, this heat exchanger represents a suitable candidate for the presented controller tuning strategy.

@@

\section{Conclusions}
\label{sec:conclusion}

This paper deals with the experimental implementation of the self-tunable approximated explicit model predictive control and provides a strategy for an effective self-tuning controller design. Based on the current reference value, the tuning parameter is scaled using linear interpolation. The previously published work related to the self-tunable explicit MPC suggested tuning just based on the distance of the reference value from the system steady-state value. This paper presents a novel idea of tuning based on the size of reference step change. The self-tuning algorithm aims to compensate for the nonlinear behavior of the controlled system. The self-tuning parameter is updated when the reference changes. It is calculated as the ratio between the size of the reference change and the maximal possible size of the reference change, which is specified before operation. Another novel contribution addresses the challenging control problem of the asymmetric system behaviour by splitting the interval of tuning parameter into two ranges, while both intervals are assigned to different operating conditions. The proposed method is implemented on a laboratory heat exchanger with nonlinear and asymmetric behavior. The non-symmetry makes the plant a suitable candidate for splitting the interval of the tuning parameter. The decision criterium is negativity or positivity of reference change. When the reference changed upwards, the control input was tuned in the first part of the interval and approached the boundary controller associated with the lower bound on the penalty matrix. On the contrary, when the reference changed downwards, the control input was tuned to approach the control input from the boundary controller with the upper bound on the penalty matrix. To investigate the control results properly, the control performance was evaluated also quantitatively. Compared to the conventional control strategy handling just a single controller, the self-tunable control approach decreased the sum-of-squared control error, maximal overshoots/undershoots, and settling time. As the consqquence, the self-tunable controllers enables significant energy savings and the corresponding reduction of the carbon footprint emissions in the industrial applications.

\section*{Acknowledgments}
